---
title: "Analysis of Mental Health Act Apprehensions in Toronto Area"
subtitle: "Simple Categorical Analysis"
author: 
  - Mingjia Chen
thanks: "Code and data are available at: https://www.toronto.ca/city-government/data-research-maps/open-data/."
date: today
date-format: long
abstract: "The report will analyzes one of the dataset publically avaliable from Toronto Open Data. To have a general idea in terms of statistics of Mental Health Act Apprehensions in Toronto, the dataset is speculated by using histograms.In general, the number of apprenhsions increases by year and a decrease in the number starts in 2022.In addition, males in age cohort 25-34 are more likely to have mental health related apprehensions."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| 
#install.packages("opendatatoronto")
#install.packages("janitor")
#install.packages("tidyverse")

library(opendatatoronto)
library(janitor)
library(tidyverse)


mental_health_package <- search_packages("Mental Health Act Apprehensions")

mental_health_resources <- 
  mental_health_package %>%
  list_package_resources()

raw_analysis_data <- 
  mental_health_resources[4,] %>%
  get_resource()

analysis_data <-
  clean_names(raw_analysis_data)
```

# Introduction

The mental health apprehensions and reports hints the overall well-being of the residents in the area. This particular report studies, by examining the Mental Health Act Apprehensions data set provided by Toronto Public Data, possible sub-population of the Toronto residents. The data was thus analyzed by the use of R language to provide a overview of the statistics both demographic and year-related in regards of the mental health apprehensions. To elaborate, the relationships between number of reports and year, gender, and age are studied in this report.

# Data {#sec-data}

The data selected for the purpose of analysis is a package data set called Mental Health Act Apprehensions from Open data Toronto by @opendatatoronto. The data consists 21 variables, including the gender, age cohort, and year of the reports. Those three variables in particular and with the counting of the number of the cases will be used for analysis.


```{r}
#| label: year-report
#| fig-cap: Counts of Apprehend Reports in Each Year
#| echo: false


analysis_data |>
  ggplot(aes(x = report_year)) +
  geom_bar() +
  theme_minimal() + # Make the theme neater
  labs(x = "Year", y = "Number of reports") 

```

The year of the reports and numbers of reports recorded in a year is being visualized by a histogram. As we can see, the number increases as the year increases until 2022. After 2022, the number of apprehensions decreases @year-report. This might be an indicator where residents' mental healthy generally becomes better after 2022. The differences between year categories might not nesscarily be significant at this point of analysis.

```{r}
#| label: fig-group
#| fig-cap: Gender and Age of People Apprehended
#| fig-subcap: ["Gender","Age"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

analysis_data |>
  ggplot(aes(x = sex)) +
  geom_histogram(stat = "count")+
  theme_minimal() + # Make the theme neater
  labs(x = "Gender", y = "Number of reports") 

analysis_data |>
  ggplot(aes(x = age_cohort)) +
  geom_histogram(stat = "count")+
  theme_minimal() + # Make the theme neater
  labs(x = "Age Cohort", y = "Number of reports") 


```
\break
With sketching histograms, the number of reports in different groups of population are compared to each other based on gender and age. In addition, the two histograms are cross-referenced (@fig-group). As the result, males with age from 25 to 34 are more likely to report mental health apprehensions. The differences between gender group are not tested as siginificant, and the analysis has limitations as it only includes binary gender (i.e., female vs. male).


# Discussion and Conclusion

Toronto is one of the major cities in Canada, which is also the capital of Province Ontario. Toronto fields includes old 140 City of Toronto Neighbourhoods structure as well as the new 158 City of Toronto Neighbourhoods structure.The mental health of residents draw our attention of whether people are having healthy mental status at this big city.

We downloaded the Mental Health Act Apprehensions from the Toroto Open Data website [@opendatatoronto]. We cleaned and tidied the data set using the statistical programming language R [@citeR] including the tidyverse [@tidyverse] and janitor [@janitor]. We then created three graphs in order to spaculate the relationships between counts of reports and year, gender, and age cohort (@fig-group).

We found that the number of reports increases each year until 2022. This could be due to Covid-19 Pandemic. Moreover, males in age cohort from 25 to 34 are more likely to report apprehensions. The difference between gender and age groups are not necessarily significant. Closer analysis such as model and power should be conducted for better statistical understanding of the data set.


\newpage

# References

